Red Wine Quality Exploration by Nan Li
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

# Global Option by used this code
knitr::opts_chunk$set(message = FALSE , warning = FALSE , echo = FALSE)
library(ggplot2)
#install.packages('corrplot')
library(corrplot)
#install.packages('RColorBrewer')
library(RColorBrewer)
library(memisc)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
rw <- read.csv('wineQualityReds.csv')

```

This report explores the dataset contains 1,599 red wines with 11 variables on the chemical
properties of the wine.

# Univariate Plots Section
```{r echo=FALSE, summary}
str(rw)

```

Because the first column is x that is the number of red wine, remove it first.
And there are totally 12 variables including the last column(quality).

### Summary RedWine dataset
```{r echo = FALSE}
rw <- rw[-1]
summary(rw)
```

### Alcohol - Count plot
```{r echo=FALSE, Univariate_Plots1}
qplot(x = alcohol, data = rw, binwidth = 0.1,fill = I('#F74940'))
  
```

### Alcohol - Count plot by log10 x-axis
```{r echo=FALSE, warning=FALSE, Univariate_Plots2}
qplot(x = alcohol, data = rw,binwidth = 0.004, fill = I('#F74940'))+
  scale_x_log10(breaks = c(9,10,12,14))+
  scale_y_continuous(limits = c(0,80))

#table(rw$alcohol)
```

As we knew, long tail is suit for log10. It's better to understand the distribution of alcohol. The transformed alcohol distribution appears bimodal with the peaking around 9.25 and 9.8. But there is a gap at 9.5. The reason is no red wine is between 9.45 and 9.5. 

### Alcohol - Count plot for each quality
```{r echo=FALSE,  warning=FALSE, Univariate_Plots3}
qplot(x = alcohol, data = rw,binwidth = 0.004, fill = I('#F74940'))+
  scale_x_log10(breaks = c(9,10,12,14))+
  scale_y_continuous(limits = c(0,80))+
  facet_wrap(~quality)

```

From above figure, I can tell the higher alcohol content means the higher quality wine.

### Fixed.acidity and Volatile.acidity - Count plot for each quality
```{r echo=FALSE,  warning=FALSE, Univariate_Plots4}

ggplot(aes(x = fixed.acidity), data = rw)+
  geom_histogram(binwidth = 0.1)+
  scale_y_continuous()+
  facet_wrap(~quality)

ggplot(aes(x = volatile.acidity), data = rw)+
  geom_histogram(binwidth = 0.01)+
  scale_y_continuous()+
  facet_wrap(~quality)

```

If volatile acidity is too high, it will degrade the quality of red wine due to the vinegar taste. There is no any value over 1 in higher quality 7 and 8.

### Fixed.acidity/Volatile.acidity - Count plot for each quality
```{r echo=FALSE,  warning=FALSE, Univariate_Plots5}
ggplot(aes(x = fixed.acidity/volatile.acidity), data = rw)+
  geom_histogram(binwidth = 0.1)+
  scale_y_continuous()+
  facet_wrap(~quality)

ggplot(aes(x = fixed.acidity/volatile.acidity), 
       data = subset(rw,volatile.acidity<=1))+
  geom_histogram(binwidth = 0.1)+
  scale_y_continuous()+
  facet_wrap(~quality)

summary(rw$fixed.acidity)
summary(rw$volatile.acidity)
summary(rw$fixed.acidity/rw$volatile.acidity)
```

Since fixed acidity is nonvolatile, I created a variable which is a ratio of nonvolatile to volatile acidity. But even though the new variable is sitting between 10 and 40 for the higher quality, it's hard to say the ratio is related to quality.

### Critic.acid - Count plot for each quality
```{r echo=FALSE,  warning=FALSE, Univariate_Plots6}

ggplot(aes(x = citric.acid), data = rw)+
  geom_histogram(binwidth = 0.01)+
  scale_y_continuous()

table(rw$citric.acid)

ggplot(aes(x = citric.acid), data = subset(rw,citric.acid>0))+
  geom_histogram(binwidth = 0.01)+
  scale_y_continuous()
```

Since citric acid can add 'freshness' and flavor to wines, I'm wondering whether citric acid is 0.49 which is the most freshness. And also the number of 0.00 citric acid is the most one. I'm guessing that is the missing value.

### Residual.sugar - Count plot for each quality
```{r echo=FALSE,  warning=FALSE, Univariate_Plots7}
ggplot(aes(x = residual.sugar), data = rw)+
  geom_histogram(binwidth = 0.1)

ggplot(aes(x = residual.sugar), data = rw)+
  geom_histogram(binwidth = 0.02)+
  scale_x_log10(breaks = c(2,4,6,8,10))+
  facet_wrap(~quality)

by(rw$residual.sugar,rw$quality, summary)

```
### Free.sulfur.dioxide - Count plot for each quality
```{r echo=FALSE,  warning=FALSE, Univariate_Plots8}
qplot(x = free.sulfur.dioxide, data = rw, binwidth = 1)

qplot(x = free.sulfur.dioxide, data = rw, binwidth = 0.05)+
  scale_x_log10(breaks = c(6,18,60))+
  scale_y_continuous()

qplot(x = total.sulfur.dioxide, data = rw, binwidth = 0.03)+
  scale_x_log10(breaks = c(10,30,100))+
  scale_y_continuous()+
  facet_wrap(~quality)

```

Since total sulfur dioxide includes free and bound forms of sulfur dioxide, I'm wondering whether bound form is related to the quality. I created a new variable bound form which equals total minue free sulfur dioxide.

### Total.sulfur.dioxide minus free.sulfur.dioxide - Count plot for each quality
```{r echo=FALSE,  warning=FALSE, Univariate_Plots9}
qplot(x = total.sulfur.dioxide-free.sulfur.dioxide, data = rw, binwidth = 0.05)+
  scale_y_continuous()+
  scale_x_log10()+
  facet_wrap(~quality)
```

### Density, PH and Sulphates - Count plot for each quality
```{r echo=FALSE, warning=FALSE, Univariate_Plots10}
qplot(x = density, data = rw, binwidth = 0.001)
summary(rw$density)

qplot(x = pH, data = rw)+
  facet_wrap(~quality)

qplot(x = sulphates, data = rw,binwidth = 0.01)+
  scale_x_log10()+
  facet_wrap(~quality)

```

### Quality - Count plot
```{r echo=FALSE,  warning=FALSE, Univariate_Plots11}
ggplot(aes(x = quality), data = rw)+
  geom_bar()
```


# Univariate Analysis

### What is the structure of your dataset?

In this red wine dataset, there are totally 1599 red wines with 11 input 
attributes(physicochemical) and 1 output attribute(quality). For quality, the 
range of score is from 0 to 10. 0 is the worst quality and 10 is the best quality.

### What is/are the main feature(s) of interest in your dataset?

The main features are quality, alcohol, fixed and volatile acidity, critric acid, 
free and total sulfur dioxide.

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

From most results, it's hard to make a conclude which one feature strongly effect the quality of red wine. I also think residual sugar and chlorides will effect the taste of the red wine.

### Did you create any new variables from existing variables in the dataset?

I created two new variables, one is the ratio between fixed and volatile acidity and the other one is bound sulfur dioxide which equals total minus free sulfur dioxide.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

First, I removed the x column which is index. For some features, I log-transformed
the right skewed distributions.I found many 0 values in the citric acid feature
and those are all missing values.So I decided to delete them and draw the graph.

# Bivariate Plots Section
### Correlation plot
```{r echo=FALSE,  warning=FALSE, Bivariate_Plots1}

M<-cor(rw)
M
corrplot(M, type = "upper")
```

Top 4 correlation coefficients with quality are:

alcohol = 0.4762,
volatile.acidity = -0.3906,
sulphates = 0.2514,
citric.acid = 0.2263.

The quality does not strongly correlate with any other variables. The highest correlation coefficient is 0.4762 with alcohol. So the alcohol content can decide the quality. On the other hand, the more volatile.acidity, the worse quality.

Top 4 correlation coefficients in datas are:

fixed.acidity with pH = -0.6830,
fixed.acidity with citric.acid = 0.6717,
fixed.acidity with density = 0.6680,
total.sulfur.dioxide with free.sulfur.dioxide = 0.6677.

In addition, since total.sulfur.dioxide is including free.sulfur.dioxide, they tend to correlate with each other. And fixed.acidity has a negitive coefficient with PH due to the acid. Consider the figures below.

### Fixed.acidity vs. PH
```{r echo=FALSE,  warning=FALSE, Bivariate_Plots2}
ggplot(aes(x = fixed.acidity, y = pH), data = rw)+
  geom_jitter(color = 'blue', alpha = 0.1)+
  stat_smooth(method = 'lm')
```

### Fixed.acidity vs. Citric.acid
```{r echo=FALSE,  warning=FALSE, Bivariate_Plots3}
ggplot(aes(x = fixed.acidity, y = citric.acid), data = rw)+
  geom_jitter(color = 'blue', alpha = 0.1)+
  stat_smooth(method = 'lm')
```

### Fixed.acidity vs. Density
```{r echo=FALSE,  warning=FALSE, Bivariate_Plots4}
ggplot(aes(x = fixed.acidity, y = density), data = rw)+
  geom_jitter(color = 'blue', alpha = 0.1)+
  stat_smooth(method = 'lm')
```

### Total.sulful.dioxide vs. Free.sulfur.dioxide
```{r echo=FALSE,  warning=FALSE, Bivariate_Plots5}
ggplot(aes(x = total.sulfur.dioxide, y = free.sulfur.dioxide), data = rw)+
  geom_jitter(color = 'blue', alpha = 0.1)+
  stat_smooth(method = 'lm')
```

### Alcohol vs. Quality boxplot
```{r echo=FALSE,  warning=FALSE, Bivariate_Plots6}
rw$quality <- as.factor(rw$quality) 
ggplot(aes(x = quality, y = alcohol),data = rw) +
  geom_boxplot(alpha = 0.1)+
  geom_jitter(alpha = 0.2)+
  stat_summary(fun.y = mean, geom='point',shape = 4)

by(rw$alcohol,rw$quality, summary)
```

The boxplot shows the higher quality red wine has more alcohol content.

### Sulphates vs. Quality
```{r echo=FALSE,  warning=FALSE, Bivariate_Plots7} 
ggplot(aes(x = quality, y = sulphates),data = rw) +
  geom_boxplot(alpha = 0.1)+
  geom_jitter(alpha = 0.2)+
  stat_summary(fun.y = mean, geom='point',shape = 4)

by(rw$sulphates,rw$quality, summary)
```
# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

The sulphate and alcohol contents are very important for the quality of redwine,
expecially higher quality levels such as 7 and 8.

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

For other features, it looks like fixed.acidity strongly correlated with some features.
It negatively correlated with pH because the acid leads to a lower pH. And it also
positive correlated with citric.acid and density.

### What was the strongest relationship you found?

The strongest relationship was between fixed.acidity and pH, which had a correlation
coefficient of -0.6830.

# Multivariate Plots Section
### Sulphats vs. Alcohol for different qualities
```{r echo=FALSE,  warning=FALSE, Multivariate_Plots1}

ggplot(aes(x = sulphates, y = alcohol), data = rw) + 
  geom_point(aes(color = quality),alpha = 0.5, size =2, position = 'jitter') +
  scale_color_brewer(type = 'div',
    guide = guide_legend(title = 'Clarity', reverse = T,
    override.aes = list(alpha = 1, size = 2)))#scale_color_brewer has type of color and legend
```

The higher quality red wine is tend to sit on the top right of the plot.

### Sulphates vs. Density for different qualities
```{r echo=FALSE,  warning=FALSE, Multivariate_Plots2}
ggplot(rw, aes(x = sulphates, 
                 fill=factor(quality))) + 
  geom_density(alpha=0.5) + 
  theme(legend.position=c(1,1),legend.justification=c(1,1))
```

### Linear Model Analysis
```{r echo=FALSE,  warning=FALSE, Multivariate_Plots3}
#linear model

m1<-lm(as.numeric(quality)~alcohol, data = rw)
m2<-update(m1, ~ . + sulphates)
m3<-update(m2, ~ . + volatile.acidity)
m4<-update(m3, ~ . + citric.acid)
mtable(m1,m2,m3,m4)

modelEstimate = predict(m4, rw)
estima <- data.frame(rw$quality, modelEstimate)
ggplot(aes(x = rw.quality, y = modelEstimate),data = estima)+
  geom_jitter()+
  stat_summary(fun.y = mean, geom='point',shape = 4)+
  labs(title = "Linear Model RedWine Quality Prediction",
    x = 'Real Quality', y = 'Predict Quality')+
  theme(plot.title = element_text(hjust = 0.5))


```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

Based on the graphs, the quality mainly related alcohol content and sulphates.
But it looked like the largest density is in the quality 3 or 4 of red wine.

### Were there any interesting or surprising interactions between features?

The linear model is created among four major variables. Alcohol, sulphates, volatile.acidity
and citric.acid. The coefficient of volatile.acidity is -1.265 which negatively 
effect the quality. 

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

Yes. From the figure above, I can tell the predicition of linear model is lower 
than the real quality. But the predicition quality 3 is much better. For future
work, I will try other models and compare with linear.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE,  warning=FALSE, Plot_One}

ggplot(aes(x = quality, y = alcohol),data = rw) +
  geom_boxplot(alpha = 0.1)+
  geom_jitter(alpha = 0.2)+
  stat_summary(fun.y = mean, geom='point',shape = 4)+
  labs(title = 'Quality vs. Alcohol',
       x = 'Quality',
       y = 'Alcohol (% by volume)')+
  theme(plot.title = element_text(hjust = 0.5))

```

### Description One
It shows the best redwine has a highest alcohol content. So when we find some higher
alcohol content in the redwine, it deserves to be paied.

### Plot Two
```{r echo=FALSE,  warning=FALSE, Plot_Two}
ggplot(aes(x = sulphates, y = alcohol), data = rw) + 
  geom_point(aes(color = quality),alpha = 0.5, size =2, position = 'jitter') +
  scale_color_brewer(type = 'div',
    guide = guide_legend(title = 'Clarity', reverse = T,
    override.aes = list(alpha = 1, size = 2)))+#scale_color_brewer has type of color and legend
  labs(title = 'Sulphates vs. Alcohol by Red Wine Quality',
       x = 'Sulphates (potassium sulphate - g / dm3)',
       y = 'Alcohol (% by volume)')+
  theme(plot.title = element_text(hjust = 0.5))
```

### Description Two
The trend of this figure is the higher quality is on the top right and the lower quality is on the bottom left. It's easy to conclude the alcohol and sulphates are positively correlated with quality.

### Plot Three
```{r echo=FALSE,  warning=FALSE, Plot_Three}
ggplot(aes(x = rw.quality, y = modelEstimate),data = estima)+
  geom_jitter()+
  stat_summary(fun.y = mean, geom='point',shape = 4, color = 'red')+
  labs(title = "Red Wine Quality Prediction by Linear Model",
    x = 'Real Quality', y = 'Predict Quality')+
  theme(plot.title = element_text(hjust = 0.5))
```

### Description Three
The prediction of the quality by linear model is obviously smaller than the real quality.
It reflects the linear model is not suitable for this dataset.
------

# Reflection
The red wine quality dataset contains 1599 red wines with 11 variables on the 
chemical properties of the wine. In order to explore the chemical properties 
influence the quality, I tried to describe one variable, two variables and 
multiple variables by drawing the plots. Such as the volatile acidity negatively
correlated to quality. A large amount of volatile acidity can lead to an unpleasant
vinegar taste. However, the citric acid and sulphates can improve the quality since
citric acid can add 'freshness' and flavor to wines and sulphates can contribute to
sulfur dioxide gas levels. I also created a linear model and tried to predict
the quality. The predict quality is lower than the real quality.

For data exploration, I'm thinking whether the number of dataset is a little bit small
and it is not fair enough since only three experts graded the red wine. Maybe we
can find 10 more experts to evaluate the red wine.

From the beginning to the end, I created so many plots. It is hard to tell which one 
is representative. So I wrote some comments after each plot and it will make easier
to decide which one I should put in the conclusion.

For future work, I will try other model to predict the quality and compare with
the linear model.